Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-08 prop_letter.tex
prop_letter.tex \section{Born} Updating...

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates of the quantum probability measures first, and then introduce
the Born rule which induces the quantum probability measures for each
state. When we shift our attention to QIVPMs, we also want to construct
a QIVPMs according to a simple rule. Since the ranges of the quantum
probability measures and QIVPMs are $\left[0,1\right]$ and $\mathscr{I}$,
respectively, if there is a function~$\iota$ mapping a number~$x\in\left[0,1\right]$
to $\iota\left(x\right)\in\mathscr{I}$, then we can concatenate it
with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
and get a $\mathscr{I}$-valued function~$\bar{\mu}:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$.
Although $\bar{\mu}$ and QIVPMs have the same the domain and range,
$\bar{\mu}$ is not necessary a QIVPM. For example, the following
theorem highlight a situation where $\bar{\mu}$ cannot be a QIVPM.

\begin{thm}If there exists three numbers~$x_{0}$, $x_{1}$, and
$x_{2}\in\left[0,1\right]$ such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$,
but 
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion-not}
\end{equation}
Then, there is a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
such that $\iota\circ\mu$ is not a QIVPM, where $\circ$ is the function
concatenation.\end{thm}

Consider the state
\begin{equation}
\begin{aligned}\rho={} & x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}\\
 & +\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}\textrm{ .}
\end{aligned}
\end{equation}
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray}
 &  & \begin{aligned} & \mathrel{\hphantom{=}}\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\end{aligned}
\nonumber \\
 &  & \begin{aligned} & =\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\\
 & \nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)
\end{aligned}
\\
 &  & \begin{aligned} & =\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\textrm{ ,}\end{aligned}
\nonumber 
\end{eqnarray}
and cannot be a QIVPM.

In other words, even if $\iota\circ\mu$ has the same domain and range
as a QIVPM, we may need to verify whether $\iota\circ\mu$ is a QIVPM
individually. For example, consider the following easy example. In
the four-dimensional Hilbert space, 
\begin{equation}
\iota\left(x\right)=\begin{cases}
\imposs & \textrm{, if }x=0\textrm{ ;}\\
\left[0,\tfrac{1}{2}\right] & \textrm{, if }x\in\left(0,\tfrac{1}{4}\right]\textrm{ ;}\\
\unknown & \textrm{, if }x\in\left(\tfrac{1}{4},\tfrac{3}{4}\right)\textrm{ ;}\\
\left[\tfrac{1}{2},1\right] & \textrm{, if }x\in\left[\tfrac{3}{4},1\right)\textrm{ ;}\\
\necess & \textrm{, if }x=1\textrm{ .}
\end{cases}
\end{equation}
and $\rho=\left(\proj{0}+\proj{1}+\proj{2}+\proj{3}\right)/4$ induce
a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}$ which is a
non-trivial fact~\citep{HOSTunpublished}. In order to verify whether
$\iota\circ\mu$ is a QIVPM easier, we want to impose some conditions
on $\iota$. Notice that our counterexamples are in a four-dimensional
Hilbert space, the conditions for $\iota$ for a Hilbert space of
dimension~$d\le3$ is not as strict as the general conditions.

>>>>

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates of the quantum probability measures first, and then introduce
the Born rule which induces a quantum probability measures from each
state. When we shift our attention to QIVPMs, we also want to construct
QIVPMs according to a simple rule. Since the ranges of the quantum
probability measures and QIVPMs are $\left[0,1\right]$ and $\mathscr{I}$,
respectively, if there is a function~$\iota$ mapping a number~$x\in\left[0,1\right]$
to an interval~$\iota\left(x\right)\in\mathscr{I}$, then we can
concatenate it with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
and get a $\mathscr{I}$-valued function~$\bar{\mu}:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$.
Although $\bar{\mu}$ and QIVPMs have the same the domain and range,
$\bar{\mu}$ is not necessary a QIVPM. For example, the following
theorem highlight a situation where $\bar{\mu}$ cannot be a QIVPM.

\begin{thm}Given a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$,
if there exists three numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but 
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion-not}
\end{equation}
Then, there is a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
such that $\iota\circ\mu$ is not a QIVPM, where $\circ$ is the function
concatenation.\end{thm}

Consider the state
\begin{equation}
\begin{aligned}\rho={} & x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}\\
 & +\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}\textrm{ .}
\end{aligned}
\end{equation}
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{equation}
\begin{aligned} & \mathrel{\hphantom{\nsubseteq}}\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\textrm{ ,}
\end{aligned}
\end{equation}
and cannot be a QIVPM.

In other words, even if $\iota\circ\mu$ has the same domain and range
as a QIVPM, we may need to verify whether $\iota\circ\mu$ is a QIVPM
individually. For example, in the four-dimensional Hilbert space,
\begin{equation}
\iota\left(x\right)=\begin{cases}
\imposs & \textrm{, if }x=0\textrm{ ;}\\
\left[0,\tfrac{1}{2}\right] & \textrm{, if }x\in\left(0,\tfrac{1}{4}\right]\textrm{ ;}\\
\unknown & \textrm{, if }x\in\left(\tfrac{1}{4},\tfrac{3}{4}\right)\textrm{ ;}\\
\left[\tfrac{1}{2},1\right] & \textrm{, if }x\in\left[\tfrac{3}{4},1\right)\textrm{ ;}\\
\necess & \textrm{, if }x=1\textrm{ .}
\end{cases}
\end{equation}
and $\rho=\left(\proj{0}+\proj{1}+\proj{2}+\proj{3}\right)/4$ inducing
a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}$ is a non-trivial
fact~\citep{HOSTunpublished}. In order to verify whether $\iota\circ\mu$
is a QIVPM easier, we can impose some conditions on $\iota$. Notice
that our counterexamples are in a four-dimensional Hilbert space,
the conditions for $\iota$ for a Hilbert space of dimension~$d\le3$
is not as strict as the general conditions.
